---
title: "Class 10: Structural Bioinformatics II"
author: "Sejal Patel"
format: html
---

# Comparative Analysis of ADK

ADK (Adenelate Kinase) us an imprtant drug target and we would love to know how it works - i.e. molecular mechanism

There has been lots of work done of this protein. 

# Install packages in the R console NOT your Rmd/Quarto file

install.packages("bio3d")
install.packages("devtools")
install.packages("BiocManager")

BiocManager::install("msa")
devtools::install_bitbucket("Grantlab/bio3d-view")

> Q10. Which of the packages above is found only on BioConductor and not CRAN? 

msa

> Q11. Which of the above packages is not found on BioConductor or CRAN?: 

All of them are found on Bioconductor or CRAN

> Q12. True or False? Functions from the devtools package can be used to install packages from GitHub and BitBucket?

True

```{r}
library(bio3d)
aa <- get.seq("1ake_A")

aa

```

```{r}
# Blast or hmmer search 
#b <- blast.pdb(aa)
```

```{r}
# Plot a summary of search results
#hits <- plot(b)
```

Let's see what is in our `hits` object. 

```{r}
#hits$pdb.id

hits <- NULL
hits$pdb.id <- c('1AKE_A','6S36_A','6RZE_A','3HPR_A','1E4V_A','5EJE_A','1E4Y_A','3X2S_A','6HAP_A','6HAM_A','4K46_A','3GMT_A','4PZL_A')
```

```{r}
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```

Now we can align and superimpose these structures which are all over the place. 

```{r}
# Align releated PDBs
pdbs <- pdbaln(files, fit = TRUE, exefile="msa")
```

Let's have a look at our `pdbs` object. 

```{r}
pdbs
```

Now we have out aligned and superposed structures we can perform all soens of analysid on them. Let's do PCA. 

```{r}
# Perform PCA
pc.xray <- pca(pdbs)
plot(pc.xray)
```

Results of PCA of ADK X-ray structures. Each dot represents one PDB strucutre. 

Cluster by RMSD (or any other method)

```{r}
# Calculate RMSD
rd <- rmsd(pdbs)

# Structure-based clustering
hc.rd <- hclust(dist(rd))
grps.rd <- cutree(hc.rd, k=3)

plot(pc.xray, 1:2, col=grps.rd)
```

```{r}
# Visualize first principal component
pc1 <- mktrj(pc.xray, pc=1, file="pc_1.pdb")
```

